A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-07-24 paper/elsarticle/elsarticle-template.tex
paper

Mention circles pattern: \cite{chiu2011novel}

@@

\begin{figure}
    \centering
    \subfloat[\centering Circles pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
    \qquad
    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
    \caption{Recovering 2D positions in support of projector calibration. (a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image, projected by the projector to each of the target locations in the scene plane. (b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera. A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected. Detected 2D projected circles pattern center points and chessboard corners overlain for illustration.}
    \label{fig:2d}
\end{figure}

@@

\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by \textit{projecting} a calibration pattern. The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector. We project the circles pattern image to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each. Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}). Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

\begin{figure}
    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red). Note that as in the figures that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
    \label{fig:3d}
\end{figure}

We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern. Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)). Note that in principle, we could project a chessboard pattern instead of a circles pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.

@@

\paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module. Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera. We then fed these images alongside the output of the camera calibration module to our projector calibration module. For each of the target locations, the steerable mirror was made to point to that location, the circles pattern image was projected to the scene plane, an image was using the camera, and the location was registered in the MDC-X software as a preset. The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}). Next, we warped the images to be projected to the respective locations using their corresponding homography, using a third dedicated custom module. These warped images were finally imported into the MDC-X software and associated with their respective location presets.

>>>>

Mention asymmetrical circle pattern: \cite{chiu2011novel}

@@

\begin{figure}
    \centering
    \subfloat[\centering Asymetrical circle pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
    \qquad
    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
    \caption{Recovering 2D positions in support of projector calibration. (a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image, to be projected to each of the target locations in the scene plane. (b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera. A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected. Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
    \label{fig:2d}
\end{figure}

@@

\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by \textit{projecting} a calibration pattern. The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector. We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each. Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}). Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

\begin{figure}
    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane (red). Note that as in the figures that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
    \label{fig:3d}
\end{figure}

We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern. Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)). Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.

@@

\paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module. Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera. We then fed these images alongside the output of the camera calibration module to our projector calibration module. For each of the target locations, the steerable mirror was made to point to that location, the asymmetrical circle pattern image was projected to the scene plane, an image was using the camera, and the location was registered in the MDC-X software as a preset. The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}). Next, we warped the images to be projected to the respective locations using their corresponding homography, using a third dedicated custom module. These warped images were finally imported into the MDC-X software and associated with their respective location presets.
